### 2018-04-11 paper.Rmd
finish shiny sections

The ability to easily browse and explore the CHILDES corpora is a cornerstone of the childes-db project. To this end we have created powerful yet easy-to-use interactive web applications that enable users to visualize various dimensions of the CHILDES corpus: frequency counts, mean lengths of utterance, type-token ratios, and more. All of this is doable without the requirement of understanding command-line tools or any kind of programming knowledge as had been the case with CLAN. The only true requirements are a browser and internet connection.

The interactive plots (web apps) are built using Shiny, an R package that enables easy building of web applications in R. Underneath the hood, each web app is making calls to our `childesr` API and subsequently plotting the data using the R plotting package `ggplot2`. A user's only task is to configure exactly what should be plotted through a series of buttons, sliders, and text boxes. The user may specify what collection, corpus, child, age range, caregiver, etc., should be included in a given analysis (e.g. mean length of utterance). The plot is displayed and updated in real-time, and if a user were to feel the need to explore the data independently then that data is also available for download alongside the plot. All of these analyses may also be reproduced only using the `childesr` package; the web apps were intended for the casual user who seeks to easy extract developmental indices without any technical overhead.

@@

A child's lexical statistics has long been an object of study in child language acquisition research. Namely, the frequency of a word <!-- something important about frequency and development?-->. In this web app, we provide users the ability to search for any word spoken by a participant in the CHILDES corpora and track the usage of that word by a child or caregiver over time. Because of the various toggles available to the user that can subset the data, a user may display frequency curves for a single child in the Brown corpus or all Spanish speaking children if desired. In addition, users may plot frequency curves belonging to caregivers alongside their child for convenient visual comparisons. A single word or multiple words may be entered in the input box, separated only by a comma.

@@

Characterizing lexical diversity has been another highly utli

<!-- sm: intro sentence here for the role of developmental metrics in CLA research and in downstream applications like SLP--> 
There are a number of ways to measure the lexical diversity of an individual's productive speech. <!-- sm: different intro; MLU isn't a lexical diversity measure -->
Some of these measures, including MLU-w (mean length of utterance in words), TTR (type-token ratio), MTLD (measure of textual lexical diversity), and voc-d are available in the Derived Measures app, which plots these measures across age for a given subset of data, again specified by collection, corpora, children, and speakers. <!-- sm: MLU is not a measure of lexical diversity-->
Because of the underlying gaps in available CHILDES morphological data, MLU-w is computed and not MLU-m (mean length of utterance in morphemes). <!-- sm: I think we should go ahead and compute this for transcript with MOR tiers? Given that this was one of the first uses, I don't think Brian will like the 'underlying gaps' explanation-->
In the English language, these two measures have been shown to be highly correlated. <!-- sm: citation needed -->
As with the Frequency Counts app, caregiver’s lexical diversity measures can be plotted alongside those of children.

@@

The population viewer is a tool that helps users understand the relative sizes and temporal extents of various corpora by displaying the total number of utterances or tokens in the available corpora at particular ages. In many instances the size of a given corpus may inform researchers’ decisions as to what is the best target dataset for analysis. <!-- note that the graphs don't currently include a grouping factor, so adding multiple corpora generates one line, so you can't actually compare two corpora in the same visualization -->

>>>>

The ability to easily browse and explore the CHILDES corpora is a cornerstone of the childes-db project. To this end we have created powerful yet easy-to-use interactive web applications that enable users to visualize various dimensions of the CHILDES corpus: frequency counts, mean lengths of utterance, type-token ratios, and more. All of this is doable without the requirement of understanding command-line tools or any kind of programming knowledge, as had been the case with CLAN. The only true requirements are a browser and internet connection.

The interactive plots (web apps) are built using Shiny, an R package that enables easy building of web applications in R. Underneath the hood, each web app is making calls to our `childesr` API and subsequently plotting the data using the R plotting package `ggplot2`. A user's only task is to configure exactly what should be plotted through a series of buttons, sliders, and text boxes. The user may specify what collection, corpus, child, age range, caregiver, etc., should be included in a given analysis (e.g. mean length of utterance). The plot is displayed and updated in real-time, and if a user were to want to explore the data independently then that data is also available for download alongside the plots. Every one of these analyses can also be produced using only the `childesr` package; the web apps were intended for the casual user who seeks to easily extract developmental indices without any technical overhead.

@@

A child's lexical statistics has long been an object of study in child language acquisition research. Namely, the frequency counts of various words may provide insight into the cognitive, conceptual, and linguistic experience of a young child. In this web app, we provide users the ability to search for any word spoken by a participant in the CHILDES corpora and track the usage of that word by a child or caregiver over time. Because of the various toggles available to the user that can subset the data, a user may display frequency curves for a single child in the Brown corpus or all Spanish speaking children, if desired. In addition, users may plot frequency curves belonging to caregivers alongside their child for convenient side-by-side comparisons. A single word or multiple words may be entered into the input box, separated only by a comma. <!-- cite original ChildFreq paper as inspiration? -->

@@

Researchers in child language acquisition have often assessed a child's linguistic performance by extracting various measures from their speech. These developmental metrics have also had a major application in speech-language pathology (SLP), where subpar measures may be indicative of speech, language, or communication disorders. The Derived Measures web app is built to provide the most common measures typically extracted from children's speech. The most famous of these measures, mean length of utterance (MLU), is used to characterize the linguistic productivity of children and is available through this web app. However, only MLU-w (mean length of utterance in words) is provided which has been shown to correlate with MLU-m (mean length of utterance in morphemes) in the English language <!-- citation needed -->. The other measures provided in this app are measures of lexical diversity: TTR (type-token ratio), MTLD (measure of textual lexical diversity) and HD-D (hypergeometric distribution of diversity) <!-- cite hd-d? mccarthy jarvis 2007 -->. As with the Frequency Counts app, a user may subset the data as he or she chooses, compare measures between caregivers and children, and aggregate across children from different corpora.

<!-- figure of derived measures app -->

@@

Many times a researcher will want to investigate the statistics of a given corpus (i.e the size, number of utterances, number of tokens, etc.) before choosing a target corpus for their analysis. This web app is intended to provide a basic overview of various corpora in the CHILDES ecosystem, facilitating familiarity with the different corpora and thus assisting with the overall analytic end-to-end process. <!-- note that the graphs don't currently include a grouping factor, so adding multiple corpora generates one line, so you can't actually compare two corpora in the same visualization -->

<!-- figure of population viewer app -->
